### 2019-05-13 Journal Version/2ECFDT.tex
main body changes

In Section \ref{binaryfdt} our focus was on binary MILPs. In this section, in an attempt to extend FDT to \{0,1,2\} problems we introduce an FDT algorithm for a 2-edge-connected multigraph problem. Given a graph $G=(V,E)$, a multi-subset of edges $F$ of $G$ is a 2-edge-connected multigraph of $G$ if for each set $\emptyset\subset U \subset V$, the number of edges in $F$ that have one endpoint in $U$ and one not in $U$ is at least 2. In the 2EC problem, we are given non-negative costs on the edge of $G$ and the goal is to find the minimum cost 2-edge-connected multigraph of $G$. Notice that, no optimal solution ever takes 3 copies of an edge in 2EC, hence we assume that we can take  each edge at most 2 times. The natural linear programming relaxation is $\EC(G) = \{x\in [0,2]^E: x(\delta(U))\geq 2 \mbox{ for } \emptyset \subset U \subset V\}$. Notice that $\dom(\EC(G))\cap [0,2]^{E} = \EC(G)$, since 2EC is a covering problem. We want to prove Theorem \ref{FDT2EC}.
\FDTEC*
We do not know the exact value for $g_{\EC}$, but we know $\frac{6}{5} \leq g_{\EC} \leq \frac{3}{2}$ \cite{carr-ravi,Wolsey1980}. Also, we need to remark that polyhedral version of Christofides' algorithm provides a $\frac{3}{2}$-approximation for 2EC, i.e. we already have an algorithm with $C\leq \frac{3}{2}$. However, we showed in Section \ref{experiment} that in practice the constant $C$ for the FDT algorithm for 2EC is much better than $\frac{3}{2}$ for FEPs of order 10.

@@

First, we need a branching lemma. Observe that  the following branching lemma is essentially a translation of Lemma \ref{LPClemma} for $\{0,1,2\}$ problems except for one additional clause.

\begin{restatable}{lemma}{2ECLPC}
	\label{LPC2EC}
	Given $x\in \EC(G)$, and $e\in E$ we can find in polynomial time vectors $x^0,x^1$ and $x^2$ and scalars $\gamma_0,\gamma_1$, and $\gamma_2$ such that: (i) $\gamma_0 + \gamma_1 +\gamma_2 \geq \frac{ 1}{g_{\EC}}$, (ii) $x^0,x^1,$ and $x^2$ are in  $ \EC(G)$, (iii) $x^0_e=0$, $x^1_e=1$, and $x^2_e=2$, (iv) $\gamma_0 x^0 + \gamma_1{x}^1  + \gamma_2x^2\leq {x}$, (v) for $f\in E$ with ${x}_f\geq 1$, we have $x^j_f\geq 1$ for $j=0,1,2$.
\end{restatable}

@@

		Note that each $F'_i$ contains at least one copy of every edge in any path $P_f$, except for at most one edge in the path. We will obtain 2-edge-connected multigraphs $F_1,\ldots,F_k$ of $G$ using $F'_1,\ldots,F'_k$, respectively. To obtain $F_i$ first remove all $P_f$ paths from $F'_i$. Suppose there is an edge $h$ in $P_f$ such that $\chi^{F'_i}_h=0$, this means that for any edge $p\in P_f$ such that $p\neq h$, $\chi^{F'_i}_p=2$. In this case, let $\chi^{F_i}_f=2$, i.e. add two copies of $f$ to $F_i$. If there are at least two edges $h,h'\in P_f$ with $\chi^{F'_i}_h=\chi^{F'_i}_{h'} = 1$, let $\chi^{F_i}_f=1$, i.e. add one copy of $f$ to $F_i$. If for all edge $h\in P_f$, we have $\chi^{F'_i}_h=2$, then let $\chi^{F_i}_f=2$. For $f\in E$ with $x_f<1$ we have
		\begin{equation}
		\sum_{i=1}^{k}\theta_i \chi^{F_i}_f=\sum_{i=1}^{k}\theta_i \chi^{F'_i}_f\leq g_{\EC}x'_f= g_{\EC}x_f.
		\end{equation}
		In addition for $f\in E$ with $x_f\geq 1$ we have $\chi^{F_i}_f \leq \frac{\sum_{h\in P_f}\chi^{F'_i}_h}{\ceil{\frac{2}{\epsilon}}-1}$ by construction.
		\begin{align*}
			\sum_{i=1}^{k}\theta_i \chi^{F_i}_f&\leq \sum_{i=1}^{k}\theta_i\frac{\sum_{h\in P_f}\chi^{F'_i}_h}{\ceil{\frac{2}{\epsilon}}-1}\\
			&= \frac{\sum_{h\in P_f} \sum_{i=1}^{k}\theta_i\chi^{F'_i}_h}{\ceil{\frac{2}{\epsilon}}-1}\\
			&\leq \frac{\sum_{h\in P_f} g_{\EC}x'_h}{\ceil{\frac{2}{\epsilon}}-1}\\
			&= \frac{\sum_{h\in P_f} g_{\EC}x'_h}{\ceil{\frac{2}{\epsilon}}-1}\\
			&= \frac{\ceil{\frac{2}{\epsilon}}}{\ceil{\frac{2}{\epsilon}}-1}g_{\EC}x_f
		\end{align*}
		Therefore, we have 
		\begin{equation}
		x' \geq \sum_{i\in [k]: \chi^{F_i}_e=1}\frac{\theta_i(\ceil{\frac{2}{\epsilon}}-1)}{g_{\EC}\ceil{\frac{2}{\epsilon}}}\chi^{F_i}+\sum_{i\in [k]: \chi^{F_i}_e=2}\frac{\theta_i(\ceil{\frac{2}{\epsilon}}-1)}{g_{\EC}\ceil{\frac{2}{\epsilon}}}\chi^{F_i}.
		\end{equation}
		Let $x^j = \sum_{i\in [k]: \chi^{F_i}_e=j}\frac{\theta_i(\ceil{\frac{2}{\epsilon}}-1)}{g_{\EC}\ceil{\frac{2}{\epsilon}}}\chi^{F_i}$ and $\theta_j =  \sum_{i\in [k]: \chi^{F_i}_e=j}\frac{\theta_i(\ceil{\frac{2}{\epsilon}}-1)}{g_{\EC}\ceil{\frac{2}{\epsilon}}}$ for $j=0,1,2$. It is easy to check that $x^j$ , $\theta_j$ for $j=0,1,2$ is a feasible solution to the LP above. Notice that $\sum_{j=0,1,2}\theta_j = \frac{\ceil{\frac{2}{\epsilon}}-1}{g_{\EC}\ceil{\frac{2}{\epsilon}}}$. By assumption, we have $\frac{\ceil{\frac{2}{\epsilon}}-1}{g_{\EC}\ceil{\frac{2}{\epsilon}}}\leq  \frac{1}{g_{\EC}}-\epsilon$. However, this means $2\leq 1$ which is a contradiction.
	\end{proof}
	This concludes the proof. \end{proof}

@@

\begin{lemma}\label{2ecpruning}
	The FDT algorithm for 2EC in  polynomial time produces sets $L_0,\ldots,L_t$ of pairs $x\in \EC(G)$ together with multipliers $\lambda$ with the following properties.
		(a) If $x\in L_i$, then $x_{e_j}=0$ or $x_{e_j}\geq 1$ for $j=1,\ldots,i$, (b) $\sum_{(x,\lambda)\in L_i }\lambda \geq \frac{1}{g_{\EC}^i}$, (c) $\sum_{(x,\lambda)\in L_i }\lambda x \leq x^*$, (d) $|L_i|\leq t$.
\end{lemma}
The proof is similar to Lemma \ref{prune}, but we need to use property (v) in Lemma \ref{LPC2EC} to prove that (a) also holds.
\begin{proof}
	We proceed by induction on the $i$. Define $L_0=\{(x^*,1)\}$. It is easy to check all the properties are satisfied. Now, suppose by induction we have $L_{i-1}$ for some $i=1,\ldots,t$ that satisfies all the properties. For each solution $x^\ell$ in $L_{i-1}$ apply Lemma \ref{LPC2EC} on $x^\ell$ and $e_{i}$ to obtain $x^{\ell j}$ and $\lambda_{\ell j}$ for $j=0,1,2$. Let $L'$ be the collection that contains $(x^{\ell j},\lambda_\ell \cdot \lambda_{\ell j})$ for $j=0,1,2$, when applied to all $(x^\ell,\lambda_\ell)$ in $L_{i-1}$. Similar to the proof in Lemma \ref{prune} one can check that $L_i$ satisfies properties (b), (c). We now verify property (a). Consider a solution $x^\ell$ in $L_{i-1}$. For $e\in \{e_1,\ldots,e_{i-1}\}$ if $x^\ell_e =0$, then by property (iv) in Lemma \ref{LPC2EC} we have $x^{\ell j}=0$ for $j=0,1,2$. Otherwise by induction we have $x^{\ell}_{e}\geq 1$ in which case property (v) in Lemma \ref{LPC2EC} ensures that $x^{\ell j}_e\geq 1$ for $j=0,1,2$. Also, $x^{\ell j}_{e_i}= j$, so $x^{\ell j}_{e_i}=0$ or $x^{\ell j}_{e_i}\geq 1$ for $j=0,1,2$.

	Finally, if $|L'|\geq t$ we let $L_i=L'$, otherwise apply $\prun(L')$ to obtain $L_{i}$.
\end{proof}

>>>>

In Section \ref{binaryfdt} our focus was on binary MILPs. In this section, in an attempt to extend FDT to \{0,1,2\} problems we introduce an FDT algorithm for a 2-edge-connected multigraph problem. Given a graph $G=(V,E)$, a multi-subset of edges $F$ of $G$ is a 2-edge-connected multigraph of $G$ if for each set $\emptyset\subset U \subset V$, the number of edges in $F$ that have one endpoint in $U$ and one not in $U$ is at least 2. In the 2EC problem, we are given non-negative costs on the edge of $G$ and the goal is to find the minimum cost 2-edge-connected multigraph of $G$. Notice that, no optimal solution ever takes 3 copies of an edge in 2EC, hence we assume that we can take  each edge at most 2 times. The natural linear programming relaxation is $\EC(G) = \{x\in [0,2]^E: x(\delta(U))\geq 2 \mbox{ for } \emptyset \subset U \subset V\}$. Notice that $\dom(\EC(G))\cap [0,2]^{E} = \EC(G)$, since 2EC is a covering problem. We want to prove Theorem \ref{FDT2EC}.
\FDTEC*
We do not know the exact value for $g_{\EC}$, but we know $\frac{6}{5} \leq g_{\EC} \leq \frac{3}{2}$ \cite{carr-ravi,Wolsey1980}. Also, we need to remark that the polyhedral analysis of Christofides' algorithm provides a $\frac{3}{2}$-approximation for 2EC, i.e. we already have an algorithm with $C\leq \frac{3}{2}$ \cite{Wolsey1980}. However, we will show in Section \ref{experiment} that in practice the constant $C$ for the FDT algorithm for 2EC is much better than $\frac{3}{2}$ for FEPs of order 10.

@@

First, we need a branching lemma. Observe that  the following branching lemma is essentially a translation of Lemma \ref{LPClemma} for $\{0,1,2\}$ problems except for one additional clause. 
\arash{put $\hat{x}^1$}
\begin{restatable}{lemma}{2ECLPC}
	\label{LPC2EC}
	Given $x\in \EC(G)$, and $e\in E$ we can find in polynomial time vectors $\hat{x}^0,\hat{x}^1$ and $\hat{x}^2$ and scalars $\gamma_0,\gamma_1$, and $\gamma_2$ such that: (i) $\gamma_0 + \gamma_1 +\gamma_2 \geq \frac{ 1}{g_{\EC}}$, (ii) $\hat{x}^0,\hat{x}^1,$ and $\hat{x}^2$ are in  $ \EC(G)$, (iii) $\hat{x}^0_e=0$, $\hat{x}^1_e=1$, and $\hat{x}^2_e=2$, (iv) $\gamma_0 \hat{x}^0 + \gamma_1{\hat{x}}^1  + \gamma_2\hat{x}^2\leq {x}$, (v) for $f\in E$ with ${x}_f\geq 1$, we have $\hat{x}^j_f\geq 1$ for $j=0,1,2$.
\end{restatable}

@@

		Note that each $F'_i$ contains at least one copy of every edge in any path $P_f$, except for at most one edge in the path. We will obtain 2-edge-connected multigraphs $F_1,\ldots,F_k$ of $G$ using $F'_1,\ldots,F'_k$, respectively. To obtain $F_i$ first remove all $P_f$ paths from $F'_i$. Suppose there is an edge $h$ in $P_f$ such that $\chi^{F'_i}_h=0$, this means that for any edge $p\in P_f$ such that $p\neq h$, $\chi^{F'_i}_p=2$. In this case, let $\chi^{F_i}_f=2$, i.e. add two copies of $f$ to $F_i$. If there are at least two edges $h,h'\in P_f$ with $\chi^{F'_i}_h=\chi^{F'_i}_{h'} = 1$, let $\chi^{F_i}_f=1$, i.e. add one copy of $f$ to $F_i$. If for all edges $h\in P_f$, we have $\chi^{F'_i}_h=2$, then let $\chi^{F_i}_f=2$. For $f\in E$ with $x_f<1$ we have
		\begin{equation}
		\sum_{i=1}^{k}\theta_i \chi^{F_i}_f=\sum_{i=1}^{k}\theta_i \chi^{F'_i}_f\leq g_{\EC}x'_f= g_{\EC}x_f.
		\end{equation}
		In addition for $f\in E$ with $x_f\geq 1$ we have $\chi^{F_i}_f \leq \frac{\sum_{h\in P_f}\chi^{F'_i}_h}{\ceil{\frac{2}{\epsilon}}-1}$ by construction.
		\begin{align*}
			\sum_{i=1}^{k}\theta_i \chi^{F_i}_f&\leq \sum_{i=1}^{k}\theta_i\frac{\sum_{h\in P_f}\chi^{F'_i}_h}{\ceil{\frac{2}{\epsilon}}-1}\\
			&= \frac{\sum_{h\in P_f} \sum_{i=1}^{k}\theta_i\chi^{F'_i}_h}{\ceil{\frac{2}{\epsilon}}-1}\\
			&\leq \frac{\sum_{h\in P_f} g_{\EC}x'_h}{\ceil{\frac{2}{\epsilon}}-1}\\
			&= \frac{\sum_{h\in P_f} g_{\EC}x_f}{\ceil{\frac{2}{\epsilon}}-1}\\
			&= \frac{\ceil{\frac{2}{\epsilon}}}{\ceil{\frac{2}{\epsilon}}-1}g_{\EC}x_f
		\end{align*}
		Therefore, we have 
		\begin{equation}
		x_e \geq \sum_{i\in [k]: \chi^{F_i}_e=1}\frac{\theta_i(\ceil{\frac{2}{\epsilon}}-1)}{g_{\EC}\ceil{\frac{2}{\epsilon}}}\chi_e^{F_i}+\sum_{i\in [k]: \chi^{F_i}_e=2}\frac{\theta_i(\ceil{\frac{2}{\epsilon}}-1)}{g_{\EC}\ceil{\frac{2}{\epsilon}}}\chi_e^{F_i}.
		\end{equation}
		Let $x^j$ be such that $x_e^j = \sum_{i\in [k]: \chi^{F_i}_e=j}\frac{\theta_i(\ceil{\frac{2}{\epsilon}}-1)}{g_{\EC}\ceil{\frac{2}{\epsilon}}}\chi^{F_i}$ and $\theta_j =  \sum_{i\in [k]: \chi^{F_i}_e=j}\frac{\theta_i(\ceil{\frac{2}{\epsilon}}-1)}{g_{\EC}\ceil{\frac{2}{\epsilon}}}$ for $j=0,1,2$. It is easy to check that $x^j$ , $\theta_j$ for $j=0,1,2$ is a feasible solution to the LP above. Notice that $\sum_{j=0,1,2}\theta_j = \frac{\ceil{\frac{2}{\epsilon}}-1}{g_{\EC}\ceil{\frac{2}{\epsilon}}}$. By assumption, we have $\frac{\ceil{\frac{2}{\epsilon}}-1}{g_{\EC}\ceil{\frac{2}{\epsilon}}}\leq  \frac{1}{g_{\EC}}-\epsilon$. However, this means $2\leq 1$ which is a contradiction.
	\end{proof}
	This concludes the proof. \end{proof}

@@

\begin{lemma}\label{2ecpruning}
	The FDT algorithm for 2EC in  polynomial time produces sets $L_0,\ldots,L_t$ of pairs $x\in \EC(G)$ together with multipliers $\lambda$ with the following properties.
		(a) If $x\in L_i$, then $x_{e_j}=0$ or $x_{e_j}\geq 1$ for $j=1,\ldots,i$, (b) $\sum_{(x,\lambda)\in L_i }\lambda \geq \frac{1}{g_{\EC}^i}$, (c) $\sum_{(x,\lambda)\in L_i }\lambda x \leq x^*$, (d) $|L_i|\leq t$.
\end{lemma}
The proof is similar to Lemma \ref{prune}, but we need to use property (v) in Lemma \ref{LPC2EC} to prove that (a) also holds.
\begin{proof}
	We proceed by induction on $i$. Define $L_0=\{(x^*,1)\}$. It is easy to check all the properties are satisfied. Now, suppose by induction we have $L_{i-1}$ for some $i=1,\ldots,t$ that satisfies all the properties. For each solution $x^\ell$ in $L_{i-1}$ apply Lemma \ref{LPC2EC} on $x^\ell$ and $e_{i}$ to obtain $x^{\ell j}$ and $\lambda_{\ell j}$ for $j=0,1,2$. Let $L'$ be the collection that contains $(x^{\ell j},\lambda_\ell \cdot \lambda_{\ell j})$ for $j=0,1,2$, when applied to all $(x^\ell,\lambda_\ell)$ in $L_{i-1}$. Similar to the proof in Lemma \ref{prune} one can check that $L_i$ satisfies properties (b), (c). We now verify property (a). Consider a solution $x^\ell$ in $L_{i-1}$. For $e\in \{e_1,\ldots,e_{i-1}\}$ if $x^\ell_e =0$, then by property (iv) in Lemma \ref{LPC2EC} we have $x^{\ell j}_e=0$ for $j=0,1,2$. Otherwise by induction we have $x^{\ell}_{e}\geq 1$ in which case property (v) in Lemma \ref{LPC2EC} ensures that $x^{\ell j}_e\geq 1$ for $j=0,1,2$. Also, $x^{\ell j}_{e_i}= j$, so $x^{\ell j}_{e_i}=0$ or $x^{\ell j}_{e_i}\geq 1$ for $j=0,1,2$.

	Finally, if $|L'|\leq t$ we let $L_i=L'$, otherwise apply $\prun(L')$ to obtain $L_{i}$ with $|L_i|\leq t$	.
\end{proof}
